The kinematics of the Local Group in a cosmological context

### 2013-03-06 paper/reply_referee.txt

Dear Editor,

We have appreciated the positive referee's comments and the suggestions
for improvement in the paper.

In what follows below the dashes, the referees comments are indented
and our responses are not indented.

Changes in the paper are boldfaced.

The referee raised concerns on Figures that look cut off. It seems to
be a technical problem on the side of the ApJLetter submissions
software. The raw eps figures and the pdf we produce (latex, dvips,
ps2pdf) look complete (see:
https://github.com/forero/LG_Kinematics/blob/master/paper/ms.pdf ,
click on 'View Raw').  We kindly ask the Editor to pay attention to
this issue and inform us if we should make any further change to our
figures.

@@

Jaime E. Forero-Romero (on behalf of the authors)

----

	First, it would be useful if the authors were more explicit about how
	they are defining "consistency". Reading the text fairly carefully, I
	think that I was able to discern that you are using 1-sigma
	constraints on v_tan to define consistency, but this needs to be
	stated very clearly. For example, what is the quoted c.l. for the HST
	limit on V_tan listed in Table 1? This should be stated clearly in the
	table itself. In the caption of Table 3, again describe exactly how
	you are defining consistency. It would be useful to know how these
	numbers change as you very the definition (1-sigma, 2, 3?). 
	I discern from a footnote that the contours drawn in Figure 1 are
	1-sigma contours. Please state this clearly in the caption
	and/or in the text that describes the figure.

We have now included explicit mention of the 1-sigma criteria in the
abstract, Table 1, Table 3 and the caption of Figure 1. There is also
a new paragraph in Section 3.1 dicussing the results for 3-sigma
uncertainties.

@@

   	 Second, the authors make a good case pairs that have Vr/Vtan values
	 similar to those observed are somewhat rare in the simulations and
	 suggest that this might possibly imply that the LG is not typical. It
	 should be possible to ask some basic questions about the small # of
	 systems that appear consistent with the data: does it matter? Do they
	 have outlying halo spins, concentrations, or formation times? What
	 about their larger-scale environments?

The answers to these questions are being clarified in a paper under
preparation by one of the coauthors (S. Bustamante).

The largest difficulty to study the pairs consistent with Vr/Vtan is
their low number. There are only 8 of them for the general sample and
_none_ in the reduced sample. This makes any statistical claim hard to
obtain. However, we have found that these pairs prefer special large
scale environments defined by the tidal and velocity shear tensor
fields: they live in filaments. We are also investigating their
alignment with respect to the directions along this filaments. We will
include in that study the referee sugestions to study the
concentration and spin parameter.

@@

We consider that this low number matters to the extent that it shows
how unlikely is to have Local Group like kinematic configurations in
LCDM, which might be relevant to explain other features in the matter
density field such as the special alignment of satellites in the Milky
Way and Andromeda or might hint towards a constraint on the expected
MW assembly history in LCDM.

   	    Third, while this is minor in terms of content, I noticed
	    a number of	grammatical issues. I strongly suggest a
	    re-read for type-os and wording.

We have fixed all the instances noted by the referee in the report and
made a new revesion to remove other typos.

### 2013-03-06 paper/ms.tex
last version before resubmission

\begin{abstract}
Recent observations constrained the tangential velocity of M31 with
respect to the Milky Way (MW) to be $v_{\rm M31,tan}<34.4$ \kms and
the radial velocity to be in the range $v_{\rm M31,rad}=-109\pm
4.4$\kms \citep{vanderMarel12}. In this study we use a large volume
high resolution N-body cosmological simulation (Bolshoi) together with three
constrained simulations to statistically study this kinematics in the
context of the $\Lambda$CDM. The comparison of the ensembles of
simulated pairs with the observed LG {\bf at the 1-$\sigma$ level in
  the uncertainties} has been done with respect to the
observed radial and tangential velocities, the reduced orbital energy
($e_{\rm tot}$), angular momentum ($l_{\rm orb}$) and the
dimensionless spin parameter, $\lambda$. Our main results are: (i) the
preferred radial and tangential velocities for pairs in $\Lambda$CDM
are $v_{\rm r}=-80\pm 20$\kms, $v_{\rm   t}=50\pm 10$\kms, (ii) pairs
around that region are $3$ to $13$ times more common than pairs within
the observational values, (iii) $15\%$ to $24\%$ of LG-like pairs in
$\Lambda$CDM have energy and angular momentum consistent with
observations while (iv) $9\%$ to $13\%$ of pairs in the same sample
show similar values in the inferred dimensionless spin parameter. It
follows that within current observational uncertainties the
quasi-conserved quantities that characterize the orbit of
the LG, i.e. $e_{\rm tot}$, $l_{\rm   orb}$ and $\lambda$, do not
challenge the standard $\Lambda$CDM model, but the model is in tension
with regard to the actual values of the radial and tangential
velocities. This might hint to a problem of the $\Lambda$CDM model to
reproduce the observed LG.

@@

\section{Simulation and Environment}
\label{sec:methods}
\subsection{The Bolshoi and Constrained Simulations}

@@

With these parameters the mass per particle is $m_{p}=1.4\times
10^{8}$\hMsun. In this paper we use halos obtained through the
Bound Density Maxima (BDM) algorithm \citep{KlypinBDM}. The halos are
selected to have an overdensity of 200 times the critical
density. Furthermore, we only include in the analysis halos whose
center is located outside the virial radius of any other halo.  We
have obtained the data through the public available Multidark
database \footnote{{\tt http://www.multidark.org/MultiDark/}}
\citep{2011arXiv1109.0003R}. The database allows us to obtain the
comoving positions, peculiar velocities and masses for all the halos
in the simulation volume at $z=0$. The positions and velocities of
these haloes correspond to the average values of the $250$ most bound
particles. The Hubble flow is taken into account to convert the
peculiar velocities into physical velocities and allow for a
comparison with observations. We have verified that the main
conclusions of this paper hold in the case of halos defined by a FOF
algorithm with a linking length 0.17 times the interparticle
distance.

@@

The full observational characteristics that we take in this Letter for
the MW-M31 pair are listed in Table \ref{table:1}.  
A more reduced sample from the full sample has been constructed
so as to fit obey the observational bounds on the masses and
separation of the two main halos. These amount to:   
\begin{enumerate}
\item The separation between the center of mass of the halos is in the
  range $700-800$kpc \citep{ribas05,vanderMarel08}. 
\item The total mass of the two halos is in the range $1-4\times
  10^{12}$\Msun \citep{vanderMarel12}. 
\end{enumerate}

@@

\item Values in parenthesis correspond to vector
  components. $\sigma_{\bf x}$ represents the uncertainty on the
  components of vector ${\bf x}$. {\bf The uncertainties correspond to
    1-sigma values.}  
\item The values for the individual halo
masses are consistent with the priors used by \cite{vanderMarel12}.

@@

\item The value for $\log_{10}\lambda$ is obtained in this Letter
from a Monte Carlo simulation as described in Section
\ref{subsection:e-l}. 
\end{enumerate}
\label{table:1}
\end{table}

\begin{table}
\caption{Summary of results from the $\Lambda$CDM Bolshoi simulation}
\begin{center}
\begin{tabular}{cccc}\hline
 &  & Full Sample & Reduced Sample\\
$v_{\rm M31,rad}$ &(\kms) & $-70\pm 10$ & $-90\pm 10$\\
$v_{\rm M31,tan}$ &(\kms) & $50\pm 10$ & $50\pm 10$\\
$\log_{10}\lambda$& & $-1.47\pm 0.13$& $-1.34\pm 0.12$\\\hline
\end{tabular}
\end{center}
\vspace{1mm}
Note: the velocity uncertainties correspond to the minimal bin size required to
obtain robust statistics in Figure \ref{fig:rt}
\label{table:2}
\end{table}

@@

\begin{figure*}
\begin{center}
\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1a.eps}
\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1b.eps}
\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1c.eps}
\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1d.eps}
\caption{Upper panels: 2D Histograms of the radial and tangential
  velocities for LG-like halo pairs in the Bolshoi simulation.   Lower
  panels: 2D Histograms of the orbital angular momentum ($l_{\rm orb}$)
  and mechanical energy ($e_{\rm tot}$) per unit of reduced mass
  calculated considering the halos as point masses. {\bf Left (right) panels
    correspond to the full (reduced) sample}. The bar indicates
  the number of pairs in each cell. The locii for the high
  density regions in the $v_{\rm r}-v_{\rm t}$ plane are listed in
  Table \ref{table:2}.   
  The half ellipse in the $v_{\rm r}-v_{\rm t}$ plane corresponds to
  the observations {\bf with 1-$\sigma$ uncertainties}. The contour
  line in   the $e_{\rm tot}-l_{\rm orb}$ plane encloses $68\%$ of the
  Monte   Carlo generated points to estimated the uncertainties from
  the   observational values summarized in Table 1. The circles
  represent the positions of the pairs from the constrained simulations.
}
\label{fig:rt}
\end{center}

@@

\begin{table*}
\caption{Summary of the comparison of the observational results
  against $\Lambda$CDM.}
\begin{center}
\begin{tabular}{ccccc}\hline
Physical & (\%) Pairs consistent & (\%) Pairs consistent & (\%) Pairs
with highest & (\%) with highest\\ 
property & with observations (1-$\sigma$) & with
observations (1-$\sigma$)& likelihood in $\Lambda$CDM & likelihood in $\Lambda$CDM\\ 
 & (full sample) & (reduced sample) & (full sample) & (reduced sample)\\ \hline
$v_{\rm r}$-$v_{\rm t}$ & (0.4\%) 8/1923 & ($<$0.6\%)0/158 & (1\%)23/1923 & (8\%)13/158\\
$e_{\rm tot}$-$l_{\rm orb}$ & (15\%)298/1923 & (24\%)38/158 & - & -\\
$\log_{10}\lambda$ & (13\%)257/1923 & (9\%)15/158 & - & -\\
$r_{\rm t}=v_{\rm t}/v_{\rm r}$& (12\%)242/1923 & (8\%)13/158 & -& -\\\hline
\end{tabular}
\end{center}
\vspace{1mm}
Notes:\\
\begin{enumerate}
\item 
{\bf In columns 2 and 3 consistency is defined from 1-$\sigma$
  uncerainties for $v_{\rm r}$-$v_{\rm t}$, $e_{\rm tot}$-$l_{\rm
    orb}$ and $\log_{10}\lambda$. In the case of $r_{\rm t}$ is
  defined from the constrain $r_{\rm t}<0.32$ derived from the
  1$-\sigma$ observational uncertainties in the radial and tangential
  velocities.} 
\item
In columns 4 and 5 the number of pairs around the preferred region
in $v_{\rm r}-v_{\rm t}$ in $\Lambda$CDM (as shown in
Fig. \ref{fig:rt} and summarized in Table \ref{table:2}) are
calculated within the same observational 1-$\sigma$ absolute uncertainty.
\end{enumerate}
\label{table:3}
\end{table*}

%Obs (full -3sigma) 127
%Theory (full 3sigma) 130
%Obs (reduced -3sigma) 11
%Theory (reduced 3sigma) 17

@@

From these results we infer that the pairs around the preferred phase
space for $\Lambda$CDM are at least $13$ times more common than pairs
with the observed velocities for the Local group. We highlight that
this is a lower bound given that in the reduced sample none of the
pairs is found in the interval allowed by observations. The high
eccentric orbit of the observed LG  constitutes an unlikely
configurations for the $\Lambda$CDM LG-like objects. This holds for
the full and the reduced sample of objects. {\bf These
  conclusions are valid at 1-$\sigma$ level in the observational
  uncertainties. An agreement between both samples of simulated LGs and the observed LG is reached only at the 3-sigma level.}

@@

The two-body problem of point-like masses can serve as a proxy for the
dynamics of the LG and be used as a tool for  studying the LG within
the framework of the standard cosmological model. Within the model the
dynamics of the LG is governed by the (center of mass)  reduced
angular momentum $l_{\rm orb}= {{\bf r}_{\rm M31}}\times{\bf v}_{\rm M31}|$ and
reduced energy  $e_{\rm tot}=\frac{1}{2} {\bf v}_{\rm M31}|^2 - G M/|{\bf
  r}_{\rm M31}|$,  where the total mass is $M=m_{\rm M31}+m_{\rm MW}$,
the reduced mass is $\mu=m_{\rm M31}m_{\rm MW}/M$ and $G$ is the
gravitational constant.

@@

We use Monte Carlo sampling to estimate the reduced energy and angular
momentum from the observed properties in the LG listed in
Table\ref{table:1}. Figure \ref{fig:rt} also presents the results in the
plane $e_{\rm tot}-l_{\rm orb}$ following the   plotting conventions from
Figure \ref{fig:rt}. \footnote{The  1-$\sigma$ contour in the lower
  panels of Figure \ref{fig:rt}
does not cross the zero level, as can be naively expected from Figure
\ref{fig:rt}. This happens for the $\approx 80\%$ level
contour.}  As expected, the $e_{\rm tot}$  and $l_{\rm orb}$
constraints are less restrictive compared with the radial and
tangential velocity constraints; 15\% (24\%) of pairs in the full
(reduced) sample obey the 1-$\sigma$ observational constraints.
We note that for the reduced sample the preferred region is outside the
1-$\sigma$ observational contours.

All the constrained LGs of the full and reduced sample are
consistent with the 1-$\sigma$ observational uncertainty.  However a
reduction by a factor of $2$ in the uncertainty of the tangential
velocity showing that the tangential velocity is below $17$\kms would
bring these pairs and the $\Lambda$CDM expectation outside the
1-$\sigma$ observational uncertainty.

@@

Summarizing, we see a broad agreement in the total angular
momentum and energy and a marked difference with the precise balance
between today's radial and tangential velocities, {\bf i.e. head-on
  collisions of MW-like halos are not common in $\Lambda$CDM}.

Under the
approximation of conservation of the orbital angular momentum and
mechanical energy, this could only be explained if the initial
conditions for the formation of the Local Group are special in
comparison to the initial conditions of any other pair of dark matter
halos in the $\Lambda$CDM cosmology. {\bf A thorough characterization
  of the small pair sample consistent with observations in terms
  of halo properties such as concentration, spin and place in the cosmic web
  \citep{Forero-Romero09,Hoffman12} is  underway (Bustamante et al. in
  prep.)}.

{\bf The results presented in this Letter open a new window into the
  question of how unique, if at all, is the LG in a cosmological
  context. This will continue to be studied within the CLUES project
  framework.}

>>>>

\begin{abstract}
Recent observations constrained the tangential velocity of M31 with
respect to the Milky Way (MW) to be $v_{\rm M31,tan}<34.4$ \kms and
the radial velocity to be in the range $v_{\rm M31,rad}=-109\pm
4.4$\kms \citep{vanderMarel12}. In this study we use a large volume
high resolution N-body cosmological simulation (Bolshoi) together with three
constrained simulations to statistically study this kinematics in the
context of the $\Lambda$CDM. The comparison of the ensembles of
simulated pairs with the observed LG {\bf at the 1-$\sigma$ level in
  the uncertainties} has been done with respect to the
radial and tangential velocities, the reduced orbital energy
($e_{\rm tot}$), angular momentum ($l_{\rm orb}$) and the
dimensionless spin parameter, $\lambda$. Our main results are: (i) the
preferred radial and tangential velocities for pairs in $\Lambda$CDM
are $v_{\rm r}=-80\pm 20$\kms, $v_{\rm   t}=50\pm 10$\kms, (ii) pairs
around that region are $3$ to $13$ times more common than pairs within
the observational values, (iii) $15\%$ to $24\%$ of LG-like pairs in
$\Lambda$CDM have energy and angular momentum consistent with
observations while (iv) $9\%$ to $13\%$ of pairs in the same sample
show similar values in the inferred dimensionless spin parameter. It
follows that within current observational uncertainties the
quasi-conserved quantities that characterize the orbit of
the LG, i.e. $e_{\rm tot}$, $l_{\rm   orb}$ and $\lambda$, do not
challenge the standard $\Lambda$CDM model, but the model is in tension
with regard to the actual values of the radial and tangential
velocities. This might hint to a problem of the $\Lambda$CDM model to
reproduce the observed LG.

@@

\section{Simulation and Pair Samples}
\label{sec:methods}
\subsection{The Bolshoi and Constrained Simulations}

@@

With these parameters the mass per particle is $m_{p}=1.4\times
10^{8}$\hMsun. In this paper we use halos obtained through the
Bound Density Maxima (BDM) algorithm \citep{KlypinBDM}. The halos are
selected to have an overdensity of 200 times the critical
density. Furthermore, we only include in the analysis halos whose
center is located outside the virial radius of any other halo.  We
have obtained the data through the public available Multidark
database \footnote{{\tt http://www.multidark.org/MultiDark/}}
\citep{2011arXiv1109.0003R}. The database allows us to obtain the
comoving positions, peculiar velocities and masses for all the halos
in the simulation volume at $z=0$. The positions and velocities of
these haloes correspond to the average values of the $250$ most bound
particles. The Hubble flow is taken into account to convert the
peculiar velocities into physical velocities and allow for a
comparison with observations. We have verified that the main
conclusions of this paper hold in the case of halos defined by a FOF
algorithm with a linking length 0.17 times the mean interparticle
distance.

@@

The full observational characteristics that we take in this Letter for
the MW-M31 pair are listed in Table \ref{table:1}.  
A more reduced sample from the full sample has been constructed
so as to obey the observational bounds on the masses and
separation of the two main halos. These amount to:   
\begin{enumerate}
\item The separation between the center of mass of the halos is in the
  range $700-800$kpc \citep{ribas05,vanderMarel08}. 
\item The total mass of the two halos is in the range $1-4\times
  10^{12}$\Msun \citep{vanderMarel12}. 
\end{enumerate}

@@

\item Values in parenthesis correspond to vector
  components. $\sigma_{\bf x}$ represents the uncertainty on the
  components of vector ${\bf x}$. {\bf The uncertainties correspond to
    1-$\sigma$ values.}  
\item The values for the individual halo
masses are consistent with the priors used by \cite{vanderMarel12}.

@@

\item The value for $\log_{10}\lambda$ is obtained in this Letter
from a Monte Carlo simulation as described in Section 3.
\end{enumerate}
\label{table:1}
\end{table}

\begin{table}
\caption{Summary of results from the $\Lambda$CDM Bolshoi simulation}
\begin{center}
\begin{tabular}{cccc}\hline
 &  & Full Sample & Reduced Sample\\
$v_{\rm M31,rad}$ &(\kms) & $-70\pm 10$ & $-90\pm 10$\\
$v_{\rm M31,tan}$ &(\kms) & $50\pm 10$ & $50\pm 10$\\
$\log_{10}\lambda$& & $-1.47\pm 0.13$& $-1.34\pm 0.12$\\\hline
\end{tabular}
\end{center}
\vspace{1mm}
Note: the velocity uncertainties correspond to the minimum bin size required to
obtain robust statistics in Figure \ref{fig:rt}
\label{table:2}
\end{table}

@@

\begin{figure*}
\begin{center}
\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1a.eps}
\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1b.eps}
\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1c.eps}
\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1d.eps}
\caption{Upper panels: 2D histograms of the radial and tangential
  velocities for LG-like halo pairs in the Bolshoi simulation.   Lower
  panels: 2D histograms of the orbital angular momentum ($l_{\rm orb}$)
  and mechanical energy ($e_{\rm tot}$) per unit of reduced mass
  calculated considering the halos as point masses. {\bf Left (right) panels
    correspond to the full (reduced) sample}. The bar indicates
  the number of pairs in each cell. The locii for the high
  density regions in the $v_{\rm r}-v_{\rm t}$ plane are listed in
  Table \ref{table:2}.   
  The half ellipse in the $v_{\rm r}-v_{\rm t}$ plane corresponds to
  the {\bf 1-$\sigma$ uncertainties} in observations. Equivalently, the contour
  line in   the $e_{\rm tot}-l_{\rm orb}$ plane encloses $68\%$ of the
  Monte   Carlo generated points from
  the   observational values summarized in Table 1. The circles
  represent the positions of the pairs from the constrained simulations.
}
\label{fig:rt}
\end{center}

@@

\begin{table*}
\caption{Summary of the comparison of the observational results
  against $\Lambda$CDM.}
\begin{center}
\begin{tabular}{ccccc}\hline
Physical & (\%) Pairs consistent & (\%) Pairs consistent & (\%) Pairs
with highest & (\%) with highest\\ 
property & with observations (1-$\sigma$) & with
observations (1-$\sigma$)& likelihood in $\Lambda$CDM & likelihood in $\Lambda$CDM\\ 
 & (full sample) & (reduced sample) & (full sample) & (reduced sample)\\ \hline
$v_{\rm r}$-$v_{\rm t}$ & (0.4\%) 8/1923 & ($<$0.6\%)0/158 & (1\%)23/1923 & (8\%)13/158\\
$e_{\rm tot}$-$l_{\rm orb}$ & (15\%)298/1923 & (24\%)38/158 & - & -\\
$\log_{10}\lambda$ & (13\%)257/1923 & (9\%)15/158 & - & -\\
$r_{\rm t}=v_{\rm t}/v_{\rm r}$& (12\%)242/1923 & (8\%)13/158 & -& -\\\hline
\end{tabular}
\end{center}
\vspace{1mm}
Notes:\\
\begin{enumerate}
\item 
{\bf In columns 2 and 3 consistency is defined from 1-$\sigma$
  uncertainties for $v_{\rm r}$-$v_{\rm t}$, $e_{\rm tot}$-$l_{\rm
    orb}$ and $\log_{10}\lambda$. In the case of $r_{\rm t}$, is
  defined from the constrain $r_{\rm t}<0.32$ derived from the
  1$-\sigma$ observational uncertainties in the radial and tangential
  velocities.} 
\item
In columns 4 and 5 the number of pairs around the preferred region
in $v_{\rm r}-v_{\rm t}$ in $\Lambda$CDM (as shown in
Fig. \ref{fig:rt} and summarized in Table \ref{table:2}) are
calculated using the same observational 1-$\sigma$ absolute
uncertainty on $v_{\rm r}$ and $v_{\rm t}$.
\end{enumerate}
\label{table:3}
\end{table*}

@@

From these results we infer that the pairs around the preferred phase
space for $\Lambda$CDM are at least $13$ times more common than pairs
with the observed velocities for the LG. We highlight that
this is a lower bound given that in the reduced sample none of the
pairs is found in the interval allowed by observations. The high
eccentric orbit of the observed LG  constitutes an unlikely
configuration for the $\Lambda$CDM LG-like objects. This holds for
the full and the reduced sample of objects. {\bf These
  conclusions are valid at 1-$\sigma$ level in the observational
  uncertainties. An equivalence in the abundance between the pairs
  around the observational values and those around the
  preferred $\Lambda$CDM velocities is reached only at
  the 4-$\sigma$ level on the observational uncertainties.}

@@

The two-body problem of point-like masses can serve as a proxy for the
dynamics of the LG and can be used as a tool for  studying the LG within
the framework of the standard cosmological model. Within the model the
dynamics of the LG is governed by the (center of mass)  reduced
angular momentum $l_{\rm orb}= |{{\bf r}_{\rm M31}}\times{\bf v}_{\rm M31}|$ and
reduced energy  $e_{\rm tot}=\frac{1}{2} {\bf v}_{\rm M31}|^2 - G M/|{\bf
  r}_{\rm M31}|$,  where the total mass is $M=m_{\rm M31}+m_{\rm MW}$,
the reduced mass is $\mu=m_{\rm M31}m_{\rm MW}/M$ and $G$ is the
gravitational constant.

@@

We use Monte Carlo sampling to estimate the reduced energy and angular
momentum from the observed properties in the LG listed in
Table\ref{table:1}. Lower panels in Figure \ref{fig:rt} also presents the results in the
plane $e_{\rm tot}-l_{\rm orb}$\footnote{The  1-$\sigma$ contour in the lower
  panels of Figure \ref{fig:rt}
does not cross the zero level, as can be naively expected from Figure
\ref{fig:rt}. This happens for the $\approx 80\%$ contour level.}.  As
expected, the $e_{\rm tot}$  and $l_{\rm orb}$ 
constraints are less restrictive compared with the radial and
tangential velocity constraints; 15\% (24\%) of pairs in the full
(reduced) sample obey the 1-$\sigma$ observational constraints.
We note that for the reduced sample the preferred region is outside the
1-$\sigma$ observational contours.

All the constrained LGs of the full and reduced sample are
consistent with the 1-$\sigma$ observational uncertainty.  However a
an hypothetical increase by a factor of $2$ in accuracy of the
tangential velocity, showing that it is below $17$\kms, would bring
these pairs and the $\Lambda$CDM expectation outside this uncertainty
region.

@@

Summarizing, we see a broad agreement in the total angular
momentum and energy and a marked difference with the precise balance
between today's radial and tangential velocities, {\bf i.e. head-on
  collisions of MW-like halos, as described by 1-$\sigma$
  uncertainties in observations, are not common in $\Lambda$CDM}.

Under the
approximation of conservation of the orbital angular momentum and
mechanical energy, this could only be explained if the initial
conditions for the formation of the Local Group are special in
comparison to the initial conditions of any other pair of dark matter
halos in the $\Lambda$CDM cosmology. {\bf Investigating this
  perspective, by a thorough characterization  of the small pair
  sample consistent with observations in terms   of halo properties
  such as concentration, spin and place in the cosmic web
  \citep{Forero-Romero09,Hoffman12} is  underway (Bustamante et al. in
  prep.)}.

{\bf The results presented in this Letter open a new window into the
  question of how unique, if at all, is the LG in a cosmological
  context. This will continue to be studied within the framework of
  the CLUES project.}
